---
title: "Linear Regression with Advertising Data"
author: "Senjuti Kundu"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    math: katex
params:
  dataset: "socexpdata.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(ggplot2)
library(tidyverse)
```

## Problem 1

Read the following csv file in R: ("https://www.statlearning.com/s/Advertising.csv"). Let the _sales_ be your response variable (y) and _TV_ be your predictor (x). Fit a linear regression model in R. Use ggplot2 to draw a scatter plot and add the fitted line to the plot.

```{r echo=FALSE, message=FALSE}
data <- read_csv(file="Advertising.csv")
reg <- lm(sales~TV,data=data)
summary(reg)

ggplot(data, aes(x=sales,y=TV)) + geom_point() + geom_smooth(method=lm, se=FALSE)
```

## Problem 2

Load the _mpgdata_ in tidyverse package. Fit the simple linear regression with _hwy_ as response and _cty_ as predictor. Visualize the result by adding the fitted line and confidence band to the scatter plot.

```{r echo=FALSE, message=FALSE}
data("mpg")
reg2 <- lm(hwy~cty,data=mpg)
ggplot(mpg, aes(x=hwy,y=cty)) + geom_point() + geom_smooth(method=lm, se=TRUE)
```

